### 2021-05-10 initial/notes.tex
stash 0512

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}

@@

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

@@

I accidentally did some extra algebra and computed to one extra order in
$\frac{s_c}{s}$, so I might as well leave it in. We will compute carefully,
noting that $\mu_4 = \frac{\cos I}{\alpha - \sin I}, \phi_4 = 0$ is exact (where
$\alpha = s/s_c$ for brevity and compatibility with \autoref{s:p2}).
\begin{align*}
    H_4 &= -\alpha \frac{\mu_4^2}{2} + \mu_4 \cos I - \sin I \sqrt{1 - \mu_4^2}
            \cos \phi_4,\\
        &= -\frac{\alpha \cos^2 I}{2(\alpha - \sin I)^2}
            + \frac{\cos^2 I}{\alpha - \sin I}
            - \sin I \frac{\sqrt{(\alpha - \sin I)^2 - \cos^2 I}}{
                \alpha - \sin I},\\
        &= \frac{1}{\p{\alpha - \sin I}^2}
            \s{-\alpha \cos^2 I + \p{\alpha - \sin I}
                \p{\cos^2 I - \sin I \sqrt{
                    (\alpha - \sin I)^2 - \cos^2 I}}},\\
        &\approx \frac{1}{\p{\alpha - \sin I}^2}
            \s{\frac{\alpha \cos^2 I}{2} -\sin I \cos^2 I - \alpha^2 \sin I
                \p{1 - \frac{2\alpha \sin I + \cos^2 I - \sin^2 I}{2\alpha^2}}
                + \sin^2 I\alpha\p{1 - \frac{2\alpha \sin I}{2\alpha^2}}},\\
        &\approx \frac{1}{\alpha^2}\p{1 + \frac{2\sin I}{\alpha}
            + \frac{3\sin^2 I}{\alpha^2}}\p{
                -\alpha^2 \sin I + 2\sin^2 I\alpha
                    + \frac{\alpha \cos^2 I}{2}
                    - \frac{(\cos^2 I + 3\sin^2 I)\sin I}{2}},\\
        &\approx -\sin I + \frac{\cos^2 I}{2\alpha}
            + \frac{1}{\alpha^2}\s{
                -3\sin^3 I - \frac{\p{\cos^2 I + 3\sin^2 I}\sin I}{2}
                    + 4\sin^3 I + \sin I \cos^2 I},\\
        &= -\sin I + \frac{\cos^2 I}{2\alpha}
            + \frac{\p{\cos^2 I - \sin^2 I}\sin I}{2\alpha^2}.
\end{align*}

@@

>>>>

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rdil}[2]{\mathrm{d}#1 / \mathrm{d}#2}
    \newcommand*{\pdil}[2]{\partial#1 / \partial#2}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\bra}[1]{\left<#1\right|}
    \newcommand*{\ket}[1]{\left|#1\right>}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \newcommand*{\uv}[1]{\hat{\bm{#1}}}

@@

\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

@@

I accidentally did some extra algebra and computed to one extra order in
$\frac{s_c}{s}$, so I might as well leave it in. We will compute carefully,
noting that $\mu_4 = \frac{\cos I}{\alpha - \sin I}, \phi_4 = 0$ is exact (where
$\alpha = s/s_c$ for brevity and compatibility with Section~\ref{s:p2}).
\begin{align*}
    H_4 &= -\alpha \frac{\mu_4^2}{2} + \mu_4 \cos I - \sin I \sqrt{1 - \mu_4^2}
            \cos \phi_4,\\
        &= -\frac{\alpha \cos^2 I}{2(\alpha - \sin I)^2}
            + \frac{\cos^2 I}{\alpha - \sin I}
            - \sin I \frac{\sqrt{(\alpha - \sin I)^2 - \cos^2 I}}{
                \alpha - \sin I},\\
        &= \frac{1}{\p{\alpha - \sin I}^2}
            \s{-\alpha \cos^2 I + \p{\alpha - \sin I}
                \p{\cos^2 I - \sin I \sqrt{
                    (\alpha - \sin I)^2 - \cos^2 I}}},\\
        &\approx \frac{1}{\p{\alpha - \sin I}^2}
            \s{\frac{\alpha \cos^2 I}{2} -\sin I \cos^2 I - \alpha^2 \sin I
                \p{1 - \frac{2\alpha \sin I + \cos^2 I - \sin^2 I}{2\alpha^2}}
                + \sin^2 I\alpha\p{1 - \frac{2\alpha \sin I}{2\alpha^2}}},\\
        &\approx \frac{1}{\alpha^2}\p{1 + \frac{2\sin I}{\alpha}
            + \frac{3\sin^2 I}{\alpha^2}}\p{
                -\alpha^2 \sin I + 2\sin^2 I\alpha
                    + \frac{\alpha \cos^2 I}{2}
                    - \frac{(\cos^2 I + 3\sin^2 I)\sin I}{2}},\\
        &\approx -\sin I + \frac{\cos^2 I}{2\alpha}
            + \frac{1}{\alpha^2}\s{
                -3\sin^3 I - \frac{\p{\cos^2 I + 3\sin^2 I}\sin I}{2}
                    + 4\sin^3 I + \sin I \cos^2 I},\\
        &= -\sin I + \frac{\cos^2 I}{2\alpha}
            + \frac{\p{\cos^2 I - \sin^2 I}\sin I}{2\alpha^2}.
\end{align*}

@@

\section{Paper Discussion Section Calculations}

\subsection{Compact Multi-Planet Systems}

\subsubsection{Our calculation}

We focus on the USP regime. In the notation of the paper, we need
\begin{align}
    1
        &\approx \frac{2n / \Omega_{\rm s} - \cos \theta
            }{\abs{g}t_{\rm s}\sin I}.
\end{align}
One tricky thing about this calculation is that $\eta$ depends on $\Omega_{\rm
s}$ which depends on $\cos \theta$ which depends on $\eta$:
\begin{align}
    \eta &= \eta_{\rm sync} \frac{n}{\Omega_{\rm s}},\\
    \frac{n}{\Omega_{\rm s}} &= \frac{1 + \cos^2\theta}{2 \cos \theta},\\
    \cos \theta &\approx \frac{\eta \cos I}{1 + \eta \sin I}.
\end{align}
This requires that
\begin{align}
    \frac{\eta}{\eta_{\rm sync}} &\approx \frac{1 + \eta \sin I
        }{2\eta \cos I},\\
    2\eta^2 \cos I &\approx \eta_{\rm sync} + \eta_{\rm sync}\eta \sin I,\\
    \eta &= \frac{\eta_{\rm sync}\sin I + \sqrt{
        \eta_{\rm sync}^2 \sin^2 I + 8\cos I \eta_{\rm sync}}}{
            4 \cos I}.
\end{align}
It proves sufficient to use the simpler form of this:
\begin{align}
    \eta &\approx \sqrt{\frac{\eta_{\rm sync}}{2\cos I}},\\
    t_{\rm s, c} &= \frac{2n / \Omega_{\rm s}}{\abs{g}\sin I},\\
        &= \frac{2}{\abs{g}\sin I\sqrt{2\eta_{\rm sync} \cos I}}.
\end{align}
Thus, we can continue to evaluate laboriously:
\begin{align}
    \frac{1}{t_{\rm s, c}} &= \abs{g}\sin I \sqrt{\frac{\eta_{\rm sync} \cos
        I}{2}},\\
    g &= \frac{3m_p}{4M_{\star}}\p{\frac{a}{a_{\rm p}}}^3 n\cos I,\\
    \frac{1}{t_{\rm s}} &= \frac{1}{4k}\frac{3k_2}{Q} \frac{M_{\star}}{m}
        \p{\frac{R}{a}}^3 n,\\
    \eta_{\rm sync} &= \s{\frac{3m_p}{4M_{\star}}\p{\frac{a}{a_{\rm p}}}^3 n
        \cos I}
        \Bigg/ \s{\frac{3k_q}{2k} \frac{M_\star}{m}\p{\frac{R}{a}}^3 n},\\
        &= \frac{m_pm}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^3
            \p{\frac{a}{R}}^3 \cos I,\\
    \frac{1}{4k}\frac{3k_2}{Q}\frac{M_\star}{m}\p{\frac{R}{a}}^3
        &= \frac{3m_{\rm p}}{4M_{\star}}\p{\frac{a}{a_{\rm p}}}^3
            \cos I \sin I \sqrt{\frac{\cos I}{2}}
            \sqrt{\frac{m_pm}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^3
                \p{\frac{a}{R}}^3 \frac{k}{k_q}\cos I},\\
        &= \frac{3}{8}\sqrt{\frac{m_{\rm p}^3m}{M_\star^4}
            \frac{a^{12}}{a_{\rm p}^9 R^3}
            \frac{k}{k_{\rm q}}}
            \sin I \cos^2 I,\\
    a^9 &= \frac{2k_2}{Q} \frac{M_\star^3}{(mm_{\rm p})^{3/2}}
        \p{Ra_{\rm p}}^{9/2}\frac{1}{\sin I \cos^2 I}
        \p{\frac{k}{k_{\rm q}}}^{-1/2},\\
% (1e-3 * (Msun / (8 Mearth))^3 * ((2 Rearth) * (0.05 AU))^(9/2))^(1/9) / (1 AU)
% = 0.033198
    a_{\rm break} &\approx 0.033 \;\mathrm{AU}
        \p{\frac{2k_2/Q}{10^{-3}}}^{1/9}
        \p{\frac{M_\star}{M_{\odot}}}^{1/3}
        \p{\frac{8M_{\oplus}}{m}}^{-1/3}
        \p{\frac{R}{2R_{\oplus}}}^{1/2}
        \p{\frac{a_{\rm p}}{0.05\;\mathrm{AU}}}^{1/2}
        \p{\frac{1}{\sin I \cos^2 I}}^{1/9},\\
% ((8 Mearth)^2 / (Msun)^2) / 2 * (1.3)^(-2) * ((0.035 AU) / (2 Rearth))^3
% = 0.011845
    \eta_{\rm sync} &= \frac{m_pm}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^3
            \p{\frac{a}{R}}^3,\\
        &= 0.012 \p{\frac{k}{k_{\rm q}}}
                \p{\frac{m}{8 M_{\oplus}}}^2
                \p{\frac{M_\star}{M_{\odot}}}^{-2}
            \p{\frac{a}{0.035 \;\mathrm{AU}}}^{3}
                \p{\frac{P_{j + 1} / P_j}{1.3}}^{-2}
                \p{\frac{R}{2R_{\oplus}}}^{-3}.
\end{align}

\subsubsection{Revised Calculation following Millholland more closely}

We consider a detailed calculation of the $\eta_{\rm sync}$ for an inner planet
of a compact, multi-planet system, with an eye towards applying it to the
Millholland runaway USP formation scenario. In this scenario, $\rdil{\uv{s}}{t}$
does not change, and is still given by
\begin{align}
    \rd{\uv{s}}{t} &= \omega_{\rm sl}\p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}},\\
    \omega_{\rm sl} \equiv \alpha &= \frac{3k_q}{2k}\frac{M_\star}{m}
        \p{\frac{R}{a}}^3 \Omega_{\rm s},\\
        &\approx n10^{-3} \p{\frac{M_{\star}}{M_{\odot}}}
            \p{\frac{m}{8 M_{\oplus}}}^{-1}
            \p{\frac{R}{2 R_{\oplus}}}^3
            \p{\frac{a}{0.035\;\mathrm{AU}}}^3
            \p{\frac{\Omega_{\rm s}}{n}}.
\end{align}
Note, for future reference, $3k_q = k_2$.

On the other hand, the evolution of $\uv{l}$ must be done more carefully, using
secular L-L theory. If we follow Millholland's formulation (which basically
includes a $J_2$ precession term from the star on top of the LL theory), we find
that the angular momentum vectors (inclination vectors) evolve by the matrix
$\bm{B}$, which has entries
\begin{align}
    B_{jj} ={}& -n_j\s{
        \frac{3}{2}J_{2\star} \p{\frac{R_\star}{a_j}}^2
        - \frac{27}{8}J_{2\star}^2\p{\frac{R_\star}{a_j}}^4
        + \frac{1}{4}\sum\limits_{k \neq j}
            \frac{M_{p, k}}{M_\star + M_{p, j}}
            \alpha_{jk} \bar{\alpha}_{jk}
            b_{3/2}^{(1)}\p{\alpha_{jk}}},\\
        \approx{}&
            -n_{j}\Bigg(\frac{3}{2}10^{-3}\frac{1}{60}
                \p{\frac{k_{2\star}}{0.2}}
                \p{\frac{P_{\star}}{\mathrm{day}}}^{-2}
                \p{\frac{R_{\star}}{R_{\odot}}}^5
                \p{\frac{M_\star}{M_{\odot}}}^{-1}
                \p{\frac{a_j}{0.035\;\mathrm{AU}}}^2\nonumber\\
            &+ \frac{1}{4} \p{2 \times 10^{-5}}
                \sum\limits_{k \neq j}
                \p{\frac{M_{p, k}}{8 M_{\oplus}}}
                \p{\frac{M_\star}{M_{\odot}}}^{-1}
                \alpha_{jk}\bar{\alpha}_{jk}b_{3/2}^{(1)}\p{\alpha_{jk}}
            \Bigg).
            &\\
    B_{jk} ={}& \frac{1}{4}n_j
            \frac{M_{p, k}}{M_\star + M_{p, j}}
            \alpha_{jk} \bar{\alpha}_{jk}
            b_{3/2}^{(1)}\p{\alpha_{jk}},\\
        ={}& n_j\frac{1}{4} \p{2 \times 10^{-5}}
                \p{\frac{M_{p, k}}{8 M_{\oplus}}}
                \p{\frac{M_\star}{M_{\odot}}}^{-1}
                \alpha_{jk}\bar{\alpha}_{jk}b_{3/2}^{(1)}\p{\alpha_{jk}}.
\end{align}
Above, $\alpha_{jk} = a_{\min} / a_{\max}$, $\bar{a}_{jk} = \min(\alpha_{jk},
1)$, and, in the case of $j < k$:
\begin{equation}
    \alpha_{jk}\bar{\alpha}_{jk}b_{3/2}^{(1)}
        \approx 3\p{\frac{a_j}{a_k}}^3 + \frac{43}{8}\p{\frac{a_j}{a_k}}^5
            + \dots
\end{equation}

For the fiducial values, the $J_2$ term is mildly dominant in the diagonal
terms, but as the Millholland paper uses $J_2 = 10^{-4}$ intsead, we see that it
is mildly subdominant, and we will ignore it here. Calling $n \equiv n_0$ the
mean motion of the innermost planet, we can express
\begin{align}
    \frac{B_{jj}}{n \p{2 \times 10^{-5}}}
        \p{\frac{m}{8 M_{\oplus}}}^{-1}
        \p{\frac{M_\star}{M_{\odot}}}
        &= -\p{\frac{P_j}{P_{j + 1}}}^{j} \sum\limits_{k \neq j}
            \frac{3}{4}\p{\frac{P_j}{P_{j+1}}}^{2(k - j)q/3},\\
    \frac{B_{jk}}{n \p{2 \times 10^{-5}}}
        \p{\frac{m}{8 M_{\oplus}}}^{-1}
        \p{\frac{M_\star}{M_{\odot}}}
        &= \p{\frac{P_j}{P_{j + 1}}}^{j}
            \frac{3}{4}\p{\frac{P_j}{P_{j+1}}}^{2|k - j|q/3}.
\end{align}
Here, $q = 3$ when $j < k$ else $2$. For instance, for a 3 planet system,
defining the period ratio $< 1$ to be $\gamma$:
\begin{equation}
    \bm{B} = n\p{1.5 \times 10^{-5}}
        \p{\frac{m}{8 M_{\oplus}}}
        \p{\frac{M_\star}{M_{\odot}}}^{-1}
        \begin{bmatrix}
            -(\gamma^2 + \gamma^4) & \gamma^2 & \gamma^{4}\\
            \gamma^{7/3} & -(\gamma^{7/3} + \gamma^3) & \gamma^3\\
            \gamma^{14/3} & \gamma^{10/3} & -\p{\gamma^{14/3} - \gamma^{10/3}}
    \end{bmatrix}
\end{equation}
Empirically, we find that the most negative eigenvalue of the matrix scales with
$\gamma^{2}$, or $\p{a / a_{\rm p}}^{3}$, indeed, as expected. Approximately the
maximum eigenvalue is $2\p{\frac{a_j}{a_{j + 1}}}^3$. Thus, we arrive at
\begin{equation}
    g_{\max} \approx \cos I 10^{-5}n\p{\frac{m}{8M_{\oplus}}}
        \p{\frac{M_{\star}}{M_{\odot}}}^{-1}\p{\frac{P_{j + 1} /
        P_j}{1.5}}^{-2}.
\end{equation}
Indeed, we thus see that $\eta_{\rm sync} \sim 10^{-2}$, as we calculated
before.

\subsection{WASP-12b}

\subsubsection{$\eta_{\rm sync}$}

First, we calculate the $g$ precession frequency correctly. Let's just use
vector notation for now, and consider the mutual precession of $\uv{l}$ and
$\uv{l}_{\rm p}$:
\begin{align}
    \rd{\uv{l}}{t} &= \omega_{\rm lp}\p{\uv{l} \cdot \uv{l}_{\rm p}}
        \p{\uv{l} \times \uv{l}_{\rm p}},\\
    \rd{\uv{l}_{\rm p}}{t} &= -\frac{L}{L_{\rm p}}\omega_{\rm lp}\p{\uv{l} \cdot
        \uv{l}_{\rm p}} \p{\uv{l} \times \uv{l}_{\rm p}}.
\end{align}
It is clear that the conserved quantity is $\bm{J} = \bm{L} + \bm{L}_{\rm p}$,
so we rewrite
\begin{align}
    \rd{\uv{l}}{t} &= \omega_{\rm lp} \cos I\p{\uv{l} \times
            \frac{\bm{L} + \bm{L}_{\rm p}}{L_{\rm p}}}\nonumber\\
        &= \omega_{\rm lp} \cos I
            \frac{J}{L_{\rm p}}\p{\uv{l} \times \uv{j}}.
\end{align}
Thus, the actual precession frequency to be used should be $g = \omega_{\rm lp}
\cos I \p{J / L_{\rm p}}$. For future reference, we also have that $\sin i /
L_{\rm p} = \sin i_{\rm p} / L = \sin I / J$, note from Fig.~\ref{fig:angles}.
\begin{figure}
    \centering
    \begin{tikzpicture}
        \filldraw (0,0) circle(1pt);
        \draw[->] (0, 0)--(-1.2, 5);
        \node[above left] at (-1.2, 5) {$L$};
        \draw[->] (0, 0)--(1.2, 0.2);
        \node[above right] at (1.2, 0.2) {$L_{\rm p}$};
        \draw[->] (0, 0)--(0, 5.2);
        \node[above] at (0, 5.2) {$J$};
        \draw[->, dashed] (-1.2, 5) -- (0, 5.2);
        \node[above] at (-0.2, 1.6) {$i$};
        \node[above right] at (0, 0) {$i_{\rm p}$};
        \node[below left] at (0, 5.2) {$i_{\rm p}$};
        \node[right] at (-1.2, 4.8) {$I$};
    \end{tikzpicture}
    \caption{Picture of angles.}\label{fig:angles}
\end{figure}

What then is $\eta_{\rm sync}$? Well, in the case of WASP-12b, $J / L_{\rm p}
\approx L / L_{\rm p} \approx (m/m_{\rm p})\sqrt{a / a_{\rm p}}$
\begin{align}
    \eta_{\rm sync} &\approx
            \frac{k}{k_q}\frac{m_pm}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^3
            \p{\frac{a}{R}}^3\frac{L}{L_{\rm p}}\cos I\nonumber\\
        &\approx
            \frac{k}{k_q}\frac{m^2}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^{7/2}
            \p{\frac{a}{R}}^3\cos I\nonumber\\
        &\approx
% 1/2 * (1.41 Mjup)^2 / (1.36 Msun)^2 * ((0.023 AU) / (1.89 Rjup))^3 * (0.023 / 0.05)^(3.5)
            \scinot{5}{-4} \cos I\p{\frac{m}{1.41M_J}}^2
                \p{\frac{M_\star}{1.36M_{\odot}}}^{-2}
                \p{\frac{a}{0.023\;\mathrm{AU}}}^{13/2}
                \p{\frac{a_{\rm p}}{0.05\;\mathrm{AU}}}^{-7/2}
                \p{\frac{R}{1.89R_{J}}}^{-3}.
\end{align}

\subsubsection{Constraint: Tidal Decay Rate}

We compute $\dot{a} / a$ for tidal dissipation in the HJ\@. Note here that
$M_\odot R_{\rm J}^5/M_J \approx 10 M_{\rm J} / M_\odot R_{\rm J}^5$, and so if
$Q \lesssim Q_\star$ then this is the dominant source of orbital decay.
According to the Millholland paper, $Q / Q_\star \approx 0.1$, and so planetary
tides can dominate if the planetary obliquity is substantial compared to the
stellar obliquity (and indeed, the stellar tidal dissipation alone struggles to
explain the boserved orbital decay). We thus obtain:
\begin{align}
    \frac{\dot{a}}{a} &= \frac{3k_2}{Q}\frac{M_\star}{m} \p{\frac{R}{a}}^5n
            \p{1 - \frac{\Omega_{\rm s}}{n}\cos \theta},\\
% 1 / (1.5 * 1e-7 * (1.36 Msun) / (1.41 Mjup) * ((1.89 Rjup) / (0.023 AU))^5 * (G * (1.36 Msun) / (0.023 AU)^3)^(1/2) * (1 yr))
        &= \frac{1}{\scinot{3.8}{6}\;\mathrm{yr}}
            \p{\frac{2k_2/Q}{10^{-6}}}
            \p{\frac{M_\star}{1.36 M_{\odot}}}^{3/2}
            \p{\frac{m}{1.41M_J}}^{-1}
            \p{\frac{R}{1.89R_j}}^5
            \p{\frac{a}{0.023\;\mathrm{AU}}}^{-13/2}
            \p{1 - \frac{\Omega_{\rm s}}{n}\cos \theta}.
\end{align}
Note that for $a = 0.038\;\mathrm{AU}$, the tidal decay rate is $\sim
\mathrm{Gyr}$. Since $\dot{a} \propto a^{-11/2}$, $a(t) \sim (C - t\dot{a} /
a)^{2/13}$ for some constant $C$, so we see that we really can't start afford to
start with the HJ any further out than $\sim 0.038\;\mathrm{AU}$

\subsection{RV Constraint}

What is the corresponding RV signal of a perturber? The RV signal in the star
is given as a velocity, which is just $n_{\rm p} \times a_{\rm p}\frac{m_{\rm
p}}{M_{\star}}$. Including the projection effect by an angle $i$, we should find
that
\begin{align}
    \text{RV semi-amplitude} &\approx
            n_{\rm p}a_{\rm p}\frac{m_{\rm p}}{M_{\star}} \sin i,\\
%(G * (1.36 Msun) / (0.05 AU))^(1/2) * (50 Mearth) / (1.36 Msun) / 2^(1/2)
        &\approx 12\;\mathrm{m/s}
            \p{\frac{M_\star}{1.36M_{\odot}}}^{-1/2}
            \p{\frac{a_{\rm p}}{0.05\;\mathrm{AU}}}^{-1/2}
            \p{\frac{m_{\rm p}}{50M_{\oplus}}}
            \frac{\sin i}{1/\sqrt{2}}
\end{align}

\subsubsection{Constraint of being in tCE2}

In order for WASP-12b to be in tCE2, we require that:

\begin{align}
    \frac{1}{t_{\rm s}} &\lesssim \abs{g} \sin I
        \sqrt{\frac{\eta_{\rm sync}\cos I}{2}},\\
    \frac{1}{4k}\frac{3k_2}{Q}\frac{M_\star}{m}\p{\frac{R}{a}}^3
        &\lesssim \frac{L}{L_{\rm p}}
            \frac{3m_{\rm p}}{4M_{\star}}\p{\frac{a}{a_{\rm p}}}^3
            \sin I \cos I\sqrt{\frac{\cos I}{2}}
            \sqrt{\frac{m^2}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^{7/2}
                \p{\frac{a}{R}}^3 \frac{k}{k_q}\cos I}\nonumber\\
        &\lesssim \frac{3m}{4M_{\star}}\p{\frac{a}{a_{\rm p}}}^{7/2}
            \sin I \cos I\sqrt{\frac{\cos I}{2}}
            \sqrt{\frac{m^2}{2M_\star^2}\p{\frac{a}{a_{\rm p}}}^{7/2}
                \p{\frac{a}{R}}^3 \frac{k}{k_q}\cos I}\nonumber\\
    \p{\frac{a_{\rm p}}{a}}^{21/4} &\lesssim
        \frac{3m}{4M_{\star}}
            \frac{4k}{3}\p{\frac{2k_2}{Q}}^{-1}
            \frac{m}{M_{\star}}
            \p{\frac{a}{R}}^3
            \sin I \cos^2 I
            \sqrt{\frac{m^2}{M_\star^2}
                \p{\frac{a}{R}}^3 \frac{k}{k_q}}\nonumber\\
% (3/4 * ((1.41 * Mjup) / (1.36 * Msun))^3 * (1/3) * 1e6 * ((0.023 AU) / (1.89 Rjup))^(9/2))^(4/21)
% = 3.3469
    \frac{a_{\rm p}}{a} &\lesssim 3.3
            \p{\frac{m}{1.41 M_J}}^{12/21}
            \p{\frac{M_\star}{1.36 M_{\odot}}}^{-12/21}
            \p{\frac{2k_2/Q}{10^{-6}}}^{-4/21}
            \p{\frac{a}{0.023\;\mathrm{AU}}}^{6/7}
            \p{\frac{R}{1.89 R_J}}^{-6/7}.
\end{align}
Indeed, this is not a very difficult constraint to satisfy. If anything, it is
harder to satisfy the angular momentum constraint, that
\begin{equation}
    \frac{L}{L_{\rm p}} = \frac{m}{m_{\rm p}}\sqrt{\frac{a}{a_{\rm p}}} \gg 1.
\end{equation}

\subsubsection{Constraint: Initial Orbital Stability}

With the constraint of being in tCE2 seeming relatively easy to satisfy, we also
require that the system be initially stable, which sets a minimum $a_{\rm p}$.
If the Hill radius criterion is used:
\begin{equation}
    a_2 - a_1 > 2\sqrt{3}\frac{a_1 + a_2}{2}\p{\frac{m_1 + m_2}{3m_0}}^{1/3},
\end{equation}
then we can calculate in the regime $m_2 \ll m_1$ to find:
\begin{equation}
    a_{\rm p} \gtrsim 0.049\;\mathrm{AU}.
\end{equation}

\subsubsection{Putting it all together}

Thus, the system must satisfy all of: (i) starts with an initial tidal decay
rate $\lesssim \;\mathrm{Gyr}$, (ii) the present-day tCE2 must be stable to
tidal dissipation, and (iii) cannot have a perturber above the RV noise
threshold. These constraints together give that: (i) the initial $a_{\rm i}
\lesssim 0.038\;\mathrm{AU}$, (ii) the perturber $a_{\rm p} \lesssim
0.076\;\mathrm{AU}$, and, if the perturber distance is maximal, (iii) $m_{\rm p}
\lesssim 80 M_{\oplus}$.

We verify that the angular momentum ratio of the perturber is indeed negligible:
$L / L_{\rm p} \approx 3$, which is marginally satisfied (and is better
satisfied if $a_{\rm p}$ is smaller).

Finally, re-evaluating $\eta_{\rm sync}$ at the start of the orbital decay for
the most optimistic initial conditions, where $a_{\rm p} = 0.049\;\mathrm{AU}$
and $a = 0.038\;\mathrm{AU}$, we find:
% 1/2 * (1.41 Mjup)^2 / (1.36 Msun)^2 * ((0.038 AU) / (1.89 Rjup))^3 * (0.038 / 0.049)^(3.5)
% 0.016
\begin{equation}
    \eta_{\rm sync} = 0.016 \cos I\p{\frac{m}{1.41M_J}}^2
            \p{\frac{M_\star}{1.36M_{\odot}}}^{-2}
            \p{\frac{a}{0.038\;\mathrm{AU}}}^{13/2}
            \p{\frac{a_{\rm p}}{0.049\;\mathrm{AU}}}^{-7/2}
            \p{\frac{R}{1.89R_{J}}}^{-3}.
\end{equation}
% 1/2 * (Mjup)^2 / (Msun)^2 * ((0.04 AU) / Rjup)^3 * (0.04 / 0.08)^(3.5)
Note, if we use Dong's values that we indeed get $\eta_{\rm sync} = 0.025$, in
agreement. Note that if the HJ was initially less inflated, the tidal decay rate
constraint implies that $R^5 / a^{13/2}$ must be a constant, thus, using a
smaller $R$ actually even \emph{further} decreases $\eta_{\rm sync}$ Thus, even
with the most optimistic parameter choices, we see that the formation
probability of WASP-12b due to weak tidal friction out of an initially isotropic
spin is quite improbable.
